\chapter{Proposal Description}\label{ch:proposal}

This chapter outlines the proposed solution developed to address the challenges of deploying and managing \gls{AI} models on heterogeneous \gls{IoT} devices. It is divided into two main sections: \textbf{the design of the solution} and \textbf{the implementation and development} of its core components.

The first section focuses on the design phase, starting with the overall system architecture, followed by the information model, system design, and sequence diagrams to illustrate system interactions. 

The second section covers the implementation phase, detailing the development of the proposed system (from now on, \gls{AURA}). Finally, it outlines the core technologies used, the documentation, and the codebase structure.

\section{Design of the Proposed Solution}\label{sec:proposal:design}

\input{chapters/proposal/design}

\clearpage

\section{Implementation and Development}\label{sec:proposal:implementation}

\input{chapters/proposal/implementation}

\clearpage

\section{Documentation System and Continuous Deployment}\label{sec:impl:documentation}

To ensure the long-term maintainability, clarity, and ease of onboarding for the \gls{AURA} platform, a dual-engine automated documentation pipeline is integrated directly into the repository. It targets the Python-based backend and edge services, as well as the TypeScript-based frontend, compiling them into a unified static website and deploying it automatically via \gls{CI/CD}.

\subsection{Backend Documentation (Sphinx)}
The platform and the \gls{IoT} Edge Device codebase are documented using inline docstrings following the NumPy/Google style conventions. These docstrings are compiled into high-level \gls{API} reference pages using \textbf{Sphinx}.

\subsection{Frontend Documentation (TypeDoc)}
For the frontend React/Next.js dashboard, documentation is generated directly from TypeScript source code using \textbf{TypeDoc}. Developers write standard JSDoc/TSDoc comments above types, hooks, \gls{API} client methods, and component definitions.

\subsection{Automated \acrshort{CI/CD} Workflow (GitHub Actions)}
To keep the published documentation synchronized with the codebase, a GitHub Actions workflow is defined in \texttt{.github/workflows/docs.yml}. The pipeline is triggered automatically on every push to the \texttt{main} branch. The workflow consists on:

\begin{enumerate}
    \item \textbf{Build Sphinx (build-sphinx)}: Spawns an Ubuntu container, installs Python, caches dependencies, and executes \texttt{make -C docs/sphinx html} to generate the platform documentation. The compiled static directory is saved as a temporary artifact.
    \item \textbf{Build TypeDoc (build-typedoc)}: Runs in a parallel container, sets up Node.js, installs frontend node modules, and builds the TypeScript references using \texttt{npx typedoc}. The output is uploaded as a second build artifact.
    \item \textbf{Assemble and Deploy (deploy)}: Once both build jobs finish successfully, the deployment container:
    \begin{enumerate}
        \item Downloads the Sphinx HTML artifact and writes it to \texttt{site/platform/}.
        \item Downloads the TypeDoc HTML artifact and writes it to \texttt{site/frontend/}.
        \item Creates a root landing page \texttt{site/index.html} containing links to both docsets and an instant \gls{HTML} refresh redirecting users to the main platform documentation page.
        \item Deploys the combined \texttt{site/} folder directly to GitHub Pages using the official \texttt{actions/deploy-pages} action.
    \end{enumerate}
\end{enumerate}
This continuous deployment guarantees that any change to the interfaces, peripheral drivers, or cloud services is immediately documented and published without manual intervention.
\todo{añadir github page}